### 2015-01-15 doc/tutorial/README.Rmd
small changes

# Tutorial

**SVMBridge** is a small framework for the controlling several SVM packages like LIBSVM, SVMperf, CVM from inside R. 
The goal is to allow for a rather easy way to do experiments with these different SVM packages. As the framework interacts with these packages by I/O, the focus is not on speed.

In the following we provide a short introduction to **SVMBridge**. An introductory example serves as a quickstart guide.
Note that our focus is on your comprehension of the basic functions and
applications. For detailed technical information and manual pages, please refer to
the package's [manual pages](http://aydindemircioglu.github.io/SVMBridge/man/). They are regularly updated and reflect the documentation of the current package.

1. [Quickstart](quickstart.md)
2. [In Depth Introduction](introduction.md)
3. [Wrappers](wrappers.md)