### 2015-09-15 paper/article.tex
first sketch of a paper

\documentclass[article]{jss}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% declarations for jss.cls %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%% almost as usual
\author{Ayd\i n Demircio\u{g}lu\\Ruhr-Universit\"at Bochum \And 
Hanna Houphouet\\Ruhr-Universit\"at Bochum \And 
        Daniel Horn\\TU Dortmund \AND
        Tobias Glasmachers\\Ruhr-Universit\"at Bochum \And 
        Bernd Bischl\\TU M\"unchen \And 
        Claus Weihs\\TU Dortmund}
\title{\pkg{SVMBridge}: Using SVM Solvers from R}

%% for pretty printing and a nice hypersummary also set:
\Plainauthor{Ayd\i n Demircio\u{g}lu, Daniel Horn} %% comma-separated
\Plaintitle{SVMBridge: Using SVM Solvers from R} %% without formatting
\Shorttitle{\pkg{SVMBridge}: Using SVM Solvers from R} %% a short title (if necessary)

%% an abstract and keywords
\Abstract{
  Most SVM Solver do not come with a wrapper in R.
  In case the SVM Solver is written in C++, it can be linked to
  a package via Rcpp. Although possible, this entails a lot of work to do,
  and is not possible, if the SVM Solver is written in other languages.
  Alternatively, one can call the SVM Solver from within R by a system command.
  SVMBridge eases this calls by providing a framework and ready wrappers
  for several SVM Solvers like LASVM, SVMperf, LLSVM and BVM/CVM.  
}
\Keywords{support vector machines, command line}
\Plainkeywords{support vector machines, command line} %% without formatting
%% at least one keyword must be supplied

%% publication information
%% NOTE: Typically, this can be left commented and will be filled out by the technical editor
%% \Volume{50}
%% \Issue{9}
%% \Month{June}
%% \Year{2012}
%% \Submitdate{2012-06-04}
%% \Acceptdate{2012-06-04}

%% The address of (at least) one author should be given
%% in the following format:
\Address{
  Ayd\i n Demircio\u{g}lu\\
  Institut f\"ur Neuroinformatik\\
  Ruhr-Universit\"at Bochum\\
  44790 Bochum
  E-mail: \email{aydin.demircioglu@ini.rub.de}\\
  URL: \url{http://www.ini.rub.de/PEOPLE/aydind/}
}
%% It is also possible to add a telephone and fax number
%% before the e-mail in the following format:
%% Telephone: +43/512/507-7103
%% Fax: +43/512/507-2851

%% for those who use Sweave please include the following line (with % symbols):
%% need no \usepackage{Sweave.sty}

%% end of declarations %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

%% include your article here, just as usual
%% Note that you should use the \pkg{}, \proglang{} and \code{} commands.

%% Note: If there is markup in \(sub)section, then it has to be escape as above.

\section{Introduction}

\subsection{Support Vector Machines}

Support Vector machines in general.

\subsection{SVM packages in R}

There are only a few SVM libraries readily available from within R.
This includes the kernlab, the e1071 and the lasvmR packages.
All of these link existing C++ sources directly from R and need
more or less extensive wrappers to allow the user direct access to
the options of the underlying SVM solver.

Other SVM solver, like the BudgetedSVM package, which contains
BSGD and LLSVM, or SVMperf cannot be called directly from R.
Instead, a system command has to be issued.
To ease this calls, the \pkg{SVMBridge} package has been developed.

\section{The SVMBridge Package}

\subsection{Wrapper}

A wrapper consists of several routines: It must provide routines
for assembling the command line call for training as well as testing,
reading and writing the model and a few general routines like searching
the binaries and printing.

Note that the search for binaries might have a huge toll on HPC clusters,
so it is adviced to specify the paths directly instead of relying on the 
automatism.

\section{Examples}

To use the SVMBridge, one usually perform the following steps:
Add the external wrapper to the SVMBridge, load the data, train a model, use the
model for prediction.

Notice that the data must be written by the SVMBridge before calling 
the corresponding SVM Solver, as there is no way to pass data via memory
to command line tools. Therefore

\subsection{Adding the Wrapper to the SVMBridge}

\subsection{Training a model}

\subsection{Predicions}

\subsection{Optimization Values}

At times, some of the training values are of interest,
e.g. when comparing different solvers, the primal value
of the SVM problem as well as the dual value might  be of interest.
These can be computed via the \code{optimizationValues()} routine.

\subsection{Performance Considerations}

We sum up the points to keep in mind when performance is of high priority:
Do not use the automatism to find the binaries, specify the paths by hand.
Do not load the data into memory, specify the path of the data when training instead.
Do not re-read the model into memory, let it on disk.

\section{Conclusion}

We provided a simple framework to attach many SVM solvers to the R Subsystem.
This allows for a systematic call of them.
By providing a wrapper, linking any SVM solver to R become much easier.

\section*{Acknowledgments}

\end{document}