### 2016-01-04 doc/tutorial/README.Rmd
snapshot

# Tutorial

**SVMBridge** is a small framework for the controlling several SVM packages like LIBSVM, SVMperf, CVM via command-line from inside R.  The goal is to allow for a rather easy way to do experiments with these different SVM packages, e.g. by replacing one variant with another one without nearly zero effort. As the framework interacts with these packages by file I/O, the focus is not on speed.

In the following we provide a short introduction to **SVMBridge**. An introductory example serves as a quickstart guide. Note that our focus is on your comprehension of the basic functions and
applications.

1. [Introduction](introduction.md)
2. [In Depth Introduction](detailed.md)
3. [Wrappers](wrappers.md)
4. [Helper functions](helpers.md)
5. [Windows How-To](windowsHowTo.md)