### 2016-01-14 doc/tutorial/wrappers.Rmd
small up

# Wrappers

**SVMBridge** comes with a couple of wrappers for different SVM packages. Here we list these and provide a few details on how to use them.

**Warning:** Currently we do not seek completeness. The **SVMBridge** is a classification-only package, so many options e.g. regarding regression etc are not supported at all. Some other options, although they can be used with classifications, might be missing too. In case you need something, you will need to adapt an existing classifier or simply write your own from scratch. If you do so, please share your results with us, so we can include these in future versions of the **SVMBridge**.

**Note:** For convenience, all the wrappers are installed along the package and can be accessed by specifying the correct path. See the introduction for examples on how to do this.

## LIBSVM

LIBSVM is practically the de-facto standard SVM solver.
The wrapper supports for now classification and RBF kernel only.

LIBSVM can be found at https://www.csie.ntu.edu.tw/~cjlin/libsvm/.

## LASVM

LASVM is very similar to LIBSVM wrapper, as both packages use the same parameters and the same model. LASVM has many more options than LIBSVM, as for our purpose we did not need them, for now these are not supported.

LASVM can be found at http://leon.bottou.org/projects/lasvm/.

## LLSVM

LLSVM is part of the BudgetedSVM package, see http://www.dabi.temple.edu/budgetedsvm/.

**Note**: The models saved by LLSVM are quite non-standard, as the transformation matrices
have to be saved instead of a list of support vectors. The wrapper includes a simple
reader, which is written in pure R. Therefore, the performance of this wrapper is rather
reduced. In case you need faster model readers/writers, you need to handle it by
yourself.

## Budgeted SGD

BSGD is part of the BudgetedSVM package, see http://www.dabi.temple.edu/budgetedsvm/.

**Note**: The models saved by LLSVM are quite non-standard, as the transformation matrices
have to be saved instead of a list of support vectors. The wrapper includes a simple
reader, which is written in pure R. Therefore, the performance of this wrapper is rather
reduced. In case you need faster model readers/writers, you need to handle it by
yourself.

## SVMperf

SVMperf has a great deal of options, so instead of providing all of these as parameter, we decided instead to give the user a choice over several 'configurations'.

SVMperf can be found at https://www.cs.cornell.edu/people/tj/svm_light/svm_perf.html.

**Note**: Please do not mix up SVMperf with SVMlight or SVMRank.

## CVM

**Note**: The source code for CVM cannot be found in the internet anymore.
Therefore we provide a copy of it in our github repository.

## BVM

Though BVM is practially the same software as CVM, sharing models and parameters and executables with CVM, the philosophy of **SVMBridge** is to have a separate wrapper for each method. Thus, though BVM has its own wrapper, all the parameters etc are the same as CVM, so please read the CVM section.

**Note**: The source code for BVM cannot be found in the internet anymore.
Therefore we provide a copy of it in our github repository.