### 2016-02-11 paper/article.tex
small fixes, add sketch of paper temporarily

\documentclass[article, shortnames]{jss}

% TODO:
% \proglang{C++}/R -> prolang{.}
% SVMBridge -> pkg{.}

\usepackage{amsmath, amssymb}
\usepackage{mathtools}

\newcommand{\sign}{\mathop{\mathrm{sign}}}
\newcommand{\N}{\mathbb{N}}
\newcommand{\R}{\mathbb{R}}
\newcommand{\Order}{\mathcal{O}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% declarations for jss.cls %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%% almost as usual
\author{Ayd\i n Demircio\u{g}lu\\Ruhr-Universit\"at Bochum \And 
Hanna Houphouet\\Ruhr-Universit\"at Bochum \And 
        Daniel Horn\\TU Dortmund \AND
        Tobias Glasmachers\\Ruhr-Universit\"at Bochum \And 
        Bernd Bischl\\TU M\"unchen \And 
        Claus Weihs\\TU Dortmund}
\title{\pkg{SVMBridge}: Using SVM Solvers from R}

%% for pretty printing and a nice hypersummary also set:
\Plainauthor{Ayd\i n Demircio\u{g}lu, Daniel Horn} %% comma-separated
\Plaintitle{SVMBridge: Using SVM Solvers from R} %% without formatting
\Shorttitle{\pkg{SVMBridge}: Using SVM Solvers from R} %% a short title (if necessary)

%% an abstract and keywords
\Abstract{
  Most SVM Solver do not come with a wrapper in  \proglang{R}.
  In case the SVM Solver is written in \proglang{C++}, it can be linked to
  a package via  Rcpp. Although possible, this entails a lot of work to do,
  and is not possible, if the SVM Solver is written in other languages.
  Alternatively, one can call the SVM Solver from within \proglang{R} by a system command.
  SVMBridge eases this calls by providing a framework and ready wrappers
  for several SVM Solvers like LASVM, SVMperf, LLSVM and BVM/CVM.  
}
\Keywords{support vector machines, command line}
\Plainkeywords{support vector machines, command line} %% without formatting
%% at least one keyword must be supplied

%% publication information
%% NOTE: Typically, this can be left commented and will be filled out by the technical editor
%% \Volume{50}
%% \Issue{9}
%% \Month{June}
%% \Year{2012}
%% \Submitdate{2012-06-04}
%% \Acceptdate{2012-06-04}

%% The address of (at least) one author should be given
%% in the following format:
\Address{
  Ayd\i n Demircio\u{g}lu, Hanna Houphouet, Tobias Glasmachers\\
  Institut f\"ur Neuroinformatik\\
  Ruhr-Universit\"at Bochum\\
  44790 Bochum\\
  Germany\\
  E-mail: \email{\{aydin.demircioglu, hanna.houphouet, tobias.glasmachers\}@ini.rub.de}\\
  URL: \url{http://www.ini.rub.de}\\
  \\

  Daniel Horn, Claus Weihs\\
  Fakult\"at Statistik\\
  Technische Universit\"at Dortmund\\
  44221 Dortmund\\
  Germany \\ 
  E-mail: \email{\{dhorn, bischl, weihs\}@statistik.tu-dortmund.de} \\
  URL: \url{https://www.statistik.tu-dortmund.de/computationalstats.html}\\
\\

  Bernd Bischl\\
Institut f\"ur Statistik \\
Ludwig-Maximilians-Universit\"at M\"unchen\\
80539 M\"unchen\\
  Germany \\
E-mail: \email{bernd.bischl@stat.uni-muenchen.de}\\
URL: \url{http://www.statistik.lmu.de/~bischl}
}

%% It is also possible to add a telephone and fax number
%% before the e-mail in the following format:
%% Telephone: +43/512/507-7103
%% Fax: +43/512/507-2851

%% for those who use Sweave please include the following line (with % symbols):
%% need no \usepackage{Sweave.sty}

%% end of declarations %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

%% include your article here, just as usual
%% Note that you should use the \pkg{}, \proglang{} and \code{} commands.

%% Note: If there is markup in \(sub)section, then it has to be escape as above.

\section{Introduction}

\subsection{Support Vector Machines}

Kernelized Support Vector Machines (SVM) \cite{cortes1995support}
are binary classifiers that use a kernel $k$ to allow for non-linear
classification decisions
$x \mapsto \sign( \langle w, \phi(x) \rangle + b)$, where $\phi$ is the
(non-linear) feature map corresponding to the kernel
$k(x, x') = \langle \phi(x), \phi(x') \rangle$ in the reproducing kernel
Hilbert space $\mathcal{H}$.
We use the RBF kernel $k(x,x') = e^{-\gamma || x-x' ||^2}$, since it
yields excellent performance and enjoys a universal approximation
property.
The SVM training problem is given by
\begin{equation}
\min_{w \in \mathcal{H}, b \in \R} \quad \frac{1}{2} ||w||^2 + C \cdot \sum_{i=1}^n \max \Big( 0, 1 - y_i \big( \langle w, \varphi(x_i) \rangle_{\mathcal{H}} + b \big) \Big),
\label{primalproblem}
\end{equation}
where $\big\{(x_1, y_1), \dots, (x_n, y_n)\big\}$ are the labeled
training points and $C > 0$ is a regularization parameter that controls
the complexity of the predictive model.

The problem, although convex, can be solved by different means,
e.g. LIBSVM employs a dual
decomposition method, while  BSGD is a primal stochastic
gradient descent solver.

\subsection[SVM packages in R]{SVM packages in \proglang{R}}

In general, nearly all SVM solvers come as binary packages,
often written in C. Therefore it is not possible to use them 
directly from within \proglang{R}. Only a few SVM libraries are readily available,
includ LIBSVM inside the well-known packages like \pkg{kernlab} (via \code{ksvm()})~\citep{kernlab}, 
and \pkg{e1071} (via \code{svm()})~\citep{dimitriadou2008misc}
as well as other less known ones, e.g. \pkg{SwarmSVM}~\citep{SwarmSVM} 
and \pkg{lasvmR}~\citep{lasvmR}.
All of these link existing \proglang{C++} sources directly from within \proglang{R} and need
more or less extensive wrappers to allow the user direct access to
the options of the underlying SVM solver.

Other SVM solvers, like the BudgetedSVM package~\citep{djuric2014budgetedsvm}, 
which contains BSGD and LLSVM, or SVMperf~\citep{joachims2009sparse}, cannot 
be called directly from  \proglang{R}.
Instead, a system command has to be issued.
To ease this calls, the \pkg{SVMBridge} package has been developed.

\section{The SVMBridge Package}

The \pkg{SVMBridge} provides a framework to incorporate 
SVM solvers by means of direct calls (via \code{system2()} commands).
This is different to \pkg{e1071} and \pkg{kernlab}, which link (and
thus copy) existing \proglang{C++} code directly inside the package.
\pkg{SVMBridge} works thus differently: Each SVM solver must provide
its own socalled {\it Wrapper}. This is a piece of R code that contains
code e.g. to allow the proper calling of the system binary of a SVM solver
as well as handling the corresponding models. Such wrapper can be
registered in the SVMBridge and henceforth can be used opaquely.

\section{General worflow}

To use the SVMBridge, one usually perform the following steps:
Make the external wrapper incl. software binaries known to the SVMBridge, 
load the data, train a model, use the model for prediction.

\subsection{Wrappers}

A wrapper is a S3 object: It must provide routines
for assembling the command line call for training as well as testing,
reading and writing the model and a few general routines like searching
the binaries and printing.
We provided default wrappers for LIBSVM, LASVM, CVM/BVM,
SVMperf, BSGD and LLSVM.

Make the SVMBridge aware of a new wrapper is
simply done by calling \code{addSVMPackage()}.
The SVMBridge comes with an easy mechanism to search 
the corresponding binaries (specified in the wrapper) to ease
the usage. To use this, one can instead call \code{findSVMSoftware}.

\subsection{Data}

Most SVM solver work with the LIBSVM sparse data format.
This format saves the labels together with all non-zero indicies of every data point, e.g.
$(0, 0, 0, 0.2 , 0, 1.4)^T$  belonging to class $3$, will be encoded as \texttt{'3 4:0.2 6:1.4'}.

Although reading these files into \proglang{R} is possible via e.g. the 
\pkg{e1071} package, most only provide pure \proglang{R} solutions,
and therefore suffer from suboptimal performance.
The SVMBridge package provides a simple sparse data format 
reading and writing, which are implemented in \proglang{C++} and
therefore can be nearly two orders of magnitudes faster than
the corresponding e1071 routines.

\subsection{Models}

Internally models contain support vectors and their coefficients.

\subsection{Training an SVM model}

Multiclass is supported, where possible.
Note that several SVM packages only support binary problems.
In these cases, training a one-vs-all machine is rather easy, if the
data is loaded into \proglang{R} first.

Reading a model is possible via the \code{readModelFromFile ()} function.
This will try to detect the format of the file by calling the 
\code{isModelFile} routine of each known wrapper.
Unluckily, several SVM solver use similar model formats, so that
a direct detection is not possible, e.g. CVM/BVM follow the LIBSVM
format, but add a comment (with '\#')about the running time to the
bottom of the model file. Apart from this change, there is no difference.
From a practical viewpoint we extended the LIBSVM model reader
to cope with this extra line, so that there is no need for an extra CVM/BVM reader.
This means that reading a CVM model will make the SVMBridge to detect a LIBSVM
model (if the LIBSVM wrapper is loaded).
In these cases, if multiple models claims ownership, a "default" model can be provided,
which will take precedence over other models. Without a default,
the model will be random.

\begin{CodeChunk}
\begin{CodeInput}
R> model = trainSVM (method = "LIBSVM", cost = 1.0, gamma = 2.0,
	trainDataFile = "./data.sparse", epsilon = 0.042, modelFile = modelFile)
\end{CodeInput}
\end{CodeChunk}

The training data can be passed on via a variable in memory or by specifying the path 
of the data set file. 
Note that there are two sets of variables: Those who belong to the SVMBridge,
like the trainDataFile variable or verbose flag, and those that are passed further
to the underlying SVM wrapper.
Not all options that are provided by the underlying SVM solver
are supported by the wrappers. As our focus lies on binary classification, because of time constraints 
we opted  to support only these options, e.g. we dropped the one-class SVDD in LIBSVM.
In case there is need for other options, it is 
easy to enhance the wrappers.
There is furthermore the subsampling option, which can be used for larger data sets.

\subsection{Testing a model}

%Predicting from a trained model is rather easy, as there are usually not many options.
Testing (or doing predictions) needs a trained SVM model as well as test data to work on.
As with training, testSVM will accept data and models either from memory or from a file.
%The same is true for the data to predict.  Again notice that data and models in memory 
%must be written to disk prior to calling the prediction binary.

\begin{CodeChunk}
\begin{CodeInput}
R> testObj = testSVM (model, testDataFile = './australian_scale', 
	readPredictions = TRUE)
\end{CodeInput}
\end{CodeChunk}

The returned object will contain meta information like testing times as well as the predictions.

\subsection{Other Considerations}

Although the SVMBridge was meant to be cross-platform, this goal
is hard to achieve. From a users perspective, it can be used on all three
major platforms (Linux, MacOS, Windows). The tests should
be executed to make sure the package works as intended.

We sum up the points to keep in mind when performance is of high priority:
Do not use the automatism to find the binaries, specify the paths by hand.
Do not load the data into memory, specify the path of the data when training instead.
Do not re-read the model into memory, let it on disk.

\section{Creating a Wrapper}

Adding your own SVM solver to the SVMBridge boils down to
writing a S3 class with several routines. In the inst folder you will find
a template that you can fill with your own code. Here we will go
through the details of adding the software package ...

The flow is as follows: The SVMBridge will call 
createTrainingArguments method of any wrapper. The wrapper
will return a string that contains all parameters that need to be
passed to the binary of the underlying SVM solver. Note that
this includes the model, prediction and training files.

\section{Conclusion}

We provided a simple framework, \pkg{SVMBridge}, to attach binary SVM solvers easily to the \proglang{R} landscape.
The \pkg{SVMBridge} does this by an easily extensible wrappers that manage the exchange of data and models
via command line.
%By providing such a wrapper, linking any SVM solver to \proglang{R} becomes much easier.

\section*{Acknowledgments}

We acknowledge support by the Mercator  \proglang{R}.search Center  \proglang{R}.hr,
under grant Pr-2013-0015 \textit{Support-Vektor-Maschinen f{\"u}r extrem gro{\ss}e Datenmengen} and
partial support by the German  \proglang{R}.search Foundation (DFG) within the Collaborative  \proglang{R}.search Centers SFB 823
\textit{Statistical modelling of nonlinear dynamic processes}, Project C2.

\bibliography{collected}

\end{document}